Effectiveness of community-based TURF-reserves in small-scale fisheries

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Expanded on TURFs and TURF-reserves close #17 close #18 close #21

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}. One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of no-take marine reserves (marine reserves from hereinafter) within Territorial Use Rights for Fisheries (TURFs) to create TURF-reserves \citep{costello_2010,afflerbach_2014,lester_2017}.

TURFs are a way of fisheries management in which a well defined group of fishes have exclusive access to an explicitly delimited portion of the ocean. TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{mccay_2017}. On the other hand, marine reserves are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools that limit or restrict fishing effort and gears and therefore aid in the recovery of marine stocks. Commonly known as TURF-Reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.

Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, marine reserve science has shown that these lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits resulting from reserves established within TURFs may be captured by the group of fishers with exclusive access. Although in theory these systems are successful \citep{costello_2010,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.

The performance of TURF-reserve systems must depend on how environmental and social factors combine and interact. This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}. There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.

@@

TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities. This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF. In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue. In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}). Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges. Oh these, 18 were originally implemented as community-based TURF-reserves. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same. Importance of evidence-based conservation mediterranean and as a movement \{guidetti_2002,ferraro_2006}. Whole paragraph here.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. These three case studies span a range of ecological and social conditions representative of different regions of Mexico. The objective of this work is twofold. First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves, which may inform similar processes in other countries. Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.

@@

>>>>

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}. One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{costello_2010,afflerbach_2014,lester_2017}.

TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean. TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks. Commonly known as TURF-reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.

Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits resulting from reserves established within TURFs should be captured exclusively by the group of fishers with exclusive access. Although in theory these systems are successful \citep{costello_2010,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.

TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact. This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}. There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.

Moreover, a new Mexican norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}). Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges. Of these, 18 were originally implemented as community-based TURF-reserves. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. These three case studies span a range of ecological and social conditions representative of different regions of Mexico. The objective of this work is twofold. First, to provide a triple bottom line evaluation of the effectiveness of community-based TURF-reserves, which may inform similar processes in other countries. Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.

@@

<!-- While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same. Importance of evidence-based conservation mediterranean and as a movement \{guidetti_2002,ferraro_2006}. Whole paragraph here. -->

@@

Before discussing our data collection methods and describing our analyses, our case studies warrant some more background. TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities. This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. However, marine reserves are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}). Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have takane advantage of this new tool is by implementing marine reserves within their TURFs. Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assambled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and each community usually hosts no more than one cooperative. Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves). Fishers then ask the government to grant legal recognition to their TURF-reserves under the name of Fish refuges following a series of studies outlined in \citet{nom}.

### 2018-12-04 docs/paper/Villasenor-Derbez.tex
Expanded on TURFs and TURF-reserves close #17 close #18 close #21

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
Micheli\(^{4}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA, USA \newline
\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
small-scale fisheries face great challenges since they tend to be hard
to monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MRs) can greatly improve the performance of coastal
fisheries and the health of the local resources
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintenance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
little empirical evidence of their effectiveness and the drivers of
their success \citep{afflerbach_2014-HP,lester_2017}.

The performance of these systems depends on how environmental and social
factors combine and interact. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, species richness, and densities of organisms
within the protected regions, climate change mitigation, and protection
from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
focused on the relationship between socioeconomic and governance
structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective. This is especially important in
social-ecological coastal systems dominated by close interactions and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
enforce if they are not legally recognized, and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Mexican Civil Society Organizations (CSOs) served as a
link between fishers and government, and created a legal framework that
solves this governance issue. In Mexico, a new norm was created in 2014
allowing fishers to request the legal recognition of community-based
reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
\citet{nom}). Fish refuges can be implemented as temporal or partial
reserves, which can protect one, some, or all resources within their
boundaries. Since 2012, old and new marine reserves have gained legal
recognition as Fishing Refuges. Oh these, 18 were originally implemented
as community-based TURF-reserves. However, their effectiveness has not
yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological
Systems (SES) framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico.
These three case studies span a range of ecological and social
conditions representative of different regions of Mexico. The objective
of this work is twofold. First, to provide a triple bottom line
evaluation of the effectiveness of community-based marine reserves,
which may inform similar processes in other countries. Second, to
evaluate the effectiveness of TURF-reserves established as Fish Refuges
in Mexico to identify opportunities where improvement or adjustment
might lead to increased effectiveness. We draw from lessons learned in
these three case studies and provide management recommendations to
maximize the effectiveness of community-based marine reserves in
small-scale fisheries in Mexico and in other regions around the world
where this tool is used to manage and rebuild their coastal fisheries.

@@

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
also collected during fish surveys. We define control sites as regions
with habitat characteristics similar to the corresponding reserves, and
that presumably had a similar probability of being selected as reserves
during the design phase. We focus our evaluation on sites where data are
available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves.

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for cooperatives with and without marine
reserves. Cooperatives incorporated in this analysis belong to larger
regional-level Cooperative Federations, and are exposed to the same
markets and institutional frameworks, making them plausible controls
\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
at the cooperative-year level, and revenues were adjusted to represent
2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.

@@

We evaluate the effect that marine reserves have had on four ecological
and two socioeconomic indicators (Table \ref{table:indicators}). Recall
that reserves were implemented to protect lobster and other benthic
invertebrates. However, we also use the available fish data to test for
associated co-benefits.

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
Micheli\(^{4}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA, USA \newline
\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
small-scale fisheries face great challenges since they tend to be hard
to monitor and enforce \citep{costello_2012}. One of the many important
ways to improve the performance of coastal fisheries and health of the
local resources is through the implementation of Territorial Use Rights
for Fisheries (TURFs) that contain no-take marine reserves within them,
thus creating TURF-reserve systems
\citep{costello_2010,afflerbach_2014,lester_2017}.

TURFs are a fisheries management tool in which a well defined group of
fishers have exclusive access to an explicitly delimited portion of the
ocean. TURFs promote a sense of stewardship and incentivise resource
users to sustainably manage their resources
\citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine
reserves (marine reserves from hereinafter) are areas where all
extractive activities are off-limits. These can be implemented to
protect biodiversity but also as fishery management tools that restrict
fishing effort and gears and therefore aid in the recovery of marine
stocks. Commonly known as TURF-reserves, the combination of these tools
increase the benefits of spatial access rights allowing the maintenance
of healthy resources \citep{afflerbach_2014,lester_2017}.

Research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access
\citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation
science has shown how marine reserves lead to increased biomass, species
richness, and abundance within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may
have a series of additional benefits like climate change mitigation,
protection from environmental variability, and fisheries benefits
\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
resulting from reserves established within TURFs are captured by the
group of fishers with exclusive access. Although in theory these systems
are successful \citep{costello_2010,smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success \citep{afflerbach_2014,lester_2017}.

TURF-reserve systems are inherently intrincate social-ecological
systems, and their effectiveness must depend on how environmental and
social factors combine and interact. This is especially important in
social-ecological coastal systems dominated by close interactions and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.
There is a growing body of literature focusing on the relations between
socioeconomic and governance structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
However, to our knowledge, no studies exist that evaluate TURF-reserves
from both a social and ecological perspective.

Moreover, a new Mexican norm was created in 2014 allowing fishers to
request the legal recognition of community-based reserves as ``Fish
Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since 2012,
old and new marine reserves have gained legal recognition as Fishing
Refuges. Of these, 18 were originally implemented as community-based
TURF-reserves. However, their effectiveness has not yet been formally
evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological
Systems (SES) framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico.
These three case studies span a range of ecological and social
conditions representative of different regions of Mexico. The objective
of this work is twofold. First, to provide a triple bottom line
evaluation of the effectiveness of community-based TURF-reserves, which
may inform similar processes in other countries. Second, to evaluate the
effectiveness of TURF-reserves established as Fish Refuges in Mexico to
identify opportunities where improvement or adjustment might lead to
increased effectiveness. We draw from lessons learned in these three
case studies and provide management recommendations to maximize the
effectiveness of community-based marine reserves in small-scale
fisheries in Mexico and in other regions around the world where this
tool is used to manage and rebuild their coastal fisheries.

@@

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, marine reserves are
difficult to enforce if they are not legally recognized, and fishers
rely on the exclusive access granted by the TURF.

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue
(\emph{i.e.} Fish Refuges; \citet{nom}). Fish refuges can be implemented
as temporal or partial reserves, which can protect one, some, or all
resources within their boundaries. One of the ways in which fishing
communities have takane advantage of this new tool is by implementing
marine reserves within their TURFs.

The common setup of community-based TURF-reserves in mexico is the
following. Fishers from a given community are assambled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and
each community usually hosts no more than one cooperative. Fishing
cooperatives interested in implementing marine reserves work with CSOs
to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves under the name of Fish refuges
following a series of studies outlined in \citet{nom}.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
also collected during fish surveys. We define control sites as regions
where: i) habitat characteristics are similar to the corresponding
reserves, ii) presumably had a similar probability of being selected as
reserves during the design phase, and iii) are located within the TURF
and therefore fishing occurs. We focus our evaluation on sites where
data are available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{stewartoaten_1986,depalma_2018}.

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for cooperatives with and without marine
reserves. Cooperatives incorporated in this analysis belong to larger
regional-level Cooperative Federations, and are exposed to the same
markets and institutional frameworks, making them plausible controls
\citep{mccay_2017,ayer_2018}. Landings and revenues were aggregated at
the cooperative-year level, and revenues were adjusted to represent 2014
values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.

@@

BACI designs and causal inference techniques have proven effective to
evaluate marine reserves, as they allow us to causally attribute
observed changes to the intervention
\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}. All sites
were surveyed annually, and at least once before implementation of the
reserves. We evaluate the effect that marine reserves have had on four
ecological and two socioeconomic indicators (Table
\ref{table:indicators}). Recall that reserves were implemented to
protect lobster and other benthic invertebrates. However, we also use
the available fish data to test for associated co-benefits.
